### 2009-12-10 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Revision of chapter 3, after the audio feedback from Prof. Puder. Also includes changes to chapter 2 (lower case Section)
- Capitalizing all the section to Section in all documents;
- Data-Centric storage is the term used in the community. Added reference;
- Changing drawn to figure
- Fact changed to Valid... (Old reference)
- Updating the Data Description Taxonomy Diagram, showing Annotation;
- Adding the description of relating system organization and type of query used on each of them.
- Fixed word agreement
- Additional description about the relationships.


git-svn-id: svn://localhost/net

The type of query mechanism used to directly or indirectly interrogate sensor
devices in a sensor network is directly related to the aforementioned locations
of the collected data. The in-network query processing is used when the
location of the collected data is a local storage site \cite{sn-storage01,
sn-storage03} within the network; a strategy typically used to maximize the
sensor's resources, while minimizing energy consumption \cite{sn-storage04}. On
the contrary, the a typical centralized query processing is used to search for
values over the collected data when the location is a centralized data
node \cite{sn-storage02}.

@@

After producing the ``sensed'' values for the observed parameters from the
environment, the sensor node is responsible for sending the raw data to a specific
data node defined by the network infrastructure, as delineated in section
\ref{sec:sn-infrastructure}. Thus, \cite{sn-storage03}
describes that the destination for the collected sensor data will be one of two
different types of storage devices: the \textbf{external storage} and the
\textbf{local storage}. In the former type, the sensor sends the
produced data to another sensor node that contains an external storage device.
In the latter the data is stored only within the memory of the device before its 
reading as a real-time data stream. On the whole, the sensor network
infrastructure dictates where the movement performed by the sensor data,
for example, \cite{sn-storage02} sees a wireless sensor networks as a many-to-one data
collection pattern, when its organization is based on single-hop or star. On
the other hand, data may also flow from one intermediate sensor node to another
until it reaches its final data sink when the data organization is structured
as a graph, tree \cite{sn-storage01, sn-storage03}.

@@

The query processing in sensor networks is a direct or indirect method of
accessing the collected data stored at data nodes, whose location is determined
by different facts described in the previous section. \cite{sn-storage03}
describes an example of an \textbf{in-network query processing} when sensor
nodes, acting as data nodes, provides real-time data stream as a response
to queries requesting the device's collected data. Likewise, if the collected
data is in use for a specific period of time, the data is cached in an external
storage device in a close-by data node. In this way, when a query is issued to
the sensor network with these characteristics, the query processor considers 
all data nodes before it returns its final response to the query. In comparison, a sensor
network whose topology is in the form of a star, the data sink is usually
located in the center of the structure, and thus, receives all
the collected data from other sensor devices within the network. For this reason, it 
can be seen as a \textbf{centralized query processing}. Factoring in that a centralized data
sink receives all the collected data from the peer nodes, this query processing
strategy is responsible for responding to indirect queries about the
produced data from the network.

Based on the literature review, there are advantages and disadvantages related to
each of the query processing approaches described with regards to the
infrastructure of the sensor network. \cite{sn-storage03} argues that the 
primary purpose of the in-network query processing is to minimize the amount
of data transmitted from sensor devices to a data sink, as sensor devices are
generally limited by energy use. Considering that sensor devices are also used
as data nodes, \cite{sn-storage04} states that this approach can potentially
flood all sensor nodes in the network, thereby decreasing its performance.
For this reason, \cite{sn-storage01} reports the use of the data-centric
storage by organizing the collected data based on its particular property
values. In this way, the query processor hits specific data nodes that contain
the types of data in the range of the requested data concluding that this approach 
decreases the network congestion and maximizes the
capacity on the sensor devices and the performance of the sensor network.

@@

Although the approach of centralized query processing maximizes the sensor
nodes capacity and minimize energy consumption, \cite{sn-storage02} explains
that this trade off imposes the creation of the unavoidable development of a
point of traffic concentration or the phenomenon of network congestion. For
this reason, \cite{sn-storage04} describes that different techniques used in 
distributed systems such as the use of master-slave data replication or 
database partitioning \cite{db-partitioning-relational} can be used to reduce
or mitigate the effects of this funneling.

@@

Data management for large scales of data require different techniques to store
and retrieve data, since it requires more computational power and data
retrieval efficient algorithms. \cite{sn-data-center} describes the creation of
a Sensor Data Center designed to store and analyze amounts of data on the range of
Terabytes to a future Petabytes of data. Its sensor network deals with not only
numerical sensor readings from transportation management, but also radar
images, used for historic purposes and data mining that performs traffic
pattern analysis. In this way, the authors report the use of parallel data
processing techniques in order to provide processing for the collected
data, and load balancing when storing data to minimize the sensor data
processing delays. For this reason, they suggest the use of concurrent control
such as snapshot isolation of the current data into separated replicated data.
Therefore, data access from users is not interrupted by the processing of
the data that is processed.

>>>>

The type of query mechanism used to directly or indirectly interrogate sensor
devices in a sensor network is directly related to the aforementioned locations
of the collected data. The in-network query processing is used when the
location of the collected data is a local storage site \cite{sn-storage01,
sn-data-centric-storage} within the network; a strategy typically used to maximize the
sensor's resources, while minimizing energy consumption \cite{sn-storage04}. On
the contrary, the a typical centralized query processing is used to search for
values over the collected data when the location is a centralized data
node \cite{sn-storage02}.

@@

After producing the ``sensed'' values for the observed parameters from the
environment, the sensor node is responsible for sending the raw data to a specific
data node defined by the network infrastructure, as delineated in Section
\ref{sec:sn-infrastructure}. Thus, \cite{sn-data-centric-storage}
describes that the destination for the collected sensor data will be one of two
different types of storage devices: the \textbf{external storage} and the
\textbf{local storage}. In the former type, the sensor sends the
produced data to another sensor node that contains an external storage device.
In the latter the data is stored only within the memory of the device before its 
reading as a real-time data stream. On the whole, the sensor network
infrastructure dictates where the movement performed by the sensor data,
for example, \cite{sn-storage02} sees a wireless sensor networks as a many-to-one data
collection pattern, when its organization is based on single-hop or star. On
the other hand, data may also flow from one intermediate sensor node to another
until it reaches its final data sink when the data organization is structured
as a graph, tree \cite{sn-storage01, sn-data-centric-storage}.

@@

The query processing in sensor networks is a direct or indirect method of
accessing the collected data stored at data nodes, whose location is determined
by different facts described in the previous section. \cite{sn-data-centric-storage}
describes an example of an \textbf{in-network query processing} when sensor
nodes, acting as data nodes, provides real-time data stream as a response
to queries requesting the device's collected data. Likewise, if the collected
data is in use for a specific period of time, the data is cached in an external
storage device in a close-by data node. In this way, when a query is issued to
the sensor network with these characteristics, the query processor considers 
all data nodes before it returns its final response to the query. In comparison, a sensor
network whose topology is in the form of a star, the data sink is usually
located in the center of the structure, and thus, receives all
the collected data from other sensor devices within the network. For this reason, it 
can be seen as a \textbf{centralized query processing}. Factoring in that a centralized data
sink receives all the collected data from the peer nodes, this query processing
strategy is responsible for responding to indirect queries about the
produced data from the network.

Based on the literature review, there are advantages and disadvantages related to
each of the query processing approaches described with regards to the
infrastructure of the sensor network. \cite{sn-data-centric-storage} argues that the 
primary purpose of the in-network query processing is to minimize the amount
of data transmitted from sensor devices to a data sink, as sensor devices are
generally limited by energy use. Considering that sensor devices are also used
as data nodes, \cite{sn-storage04} states that this approach can potentially
flood all sensor nodes in the network, thereby decreasing its performance.
For this reason, \cite{sn-storage01} reports the use of the data-centric
storage by organizing the collected data based on its particular property
values. In this way, the query processor hits specific data nodes that contain
the types of data in the range of the requested data concluding that this approach 
decreases the network congestion and maximizes the
capacity on the sensor devices and the performance of the sensor network.

@@

Although the approach of centralized query processing maximizes the sensor
nodes capacity and minimize energy consumption, \cite{sn-storage02} explains
that this trade off imposes the creation of the unavoidable development of a
point of traffic concentration or the phenomenon of network congestion. For
this reason, \cite{sn-storage04} describes that different techniques used in 
distributed systems such as the use of master-slave data replication
\cite{sn-data-center-replication-load-balance} or database partitioning
\cite{db-partitioning-relational} can be used to reduce or mitigate the
effects of this funneling.

@@

Data management for large scales of data require different techniques to store
and retrieve data, since it requires more computational power and data
retrieval efficient algorithms. \cite{sn-data-center-replication-load-balance} describes the creation of
a Sensor Data Center designed to store and analyze amounts of data on the range of
Terabytes to a future Petabytes of data. Its sensor network deals with not only
numerical sensor readings from transportation management, but also radar
images, used for historic purposes and data mining that performs traffic
pattern analysis. In this way, the authors report the use of parallel data
processing techniques in order to provide processing for the collected
data, and load balancing when storing data to minimize the sensor data
processing delays. For this reason, they suggest the use of concurrent control
such as snapshot isolation of the current data into separated replicated data.
Therefore, data access from users is not interrupted by the processing of
the data that is processed.

### 2009-12-10 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Revision of chapter 3, after the audio feedback from Prof. Puder. Also includes changes to chapter 2 (lower case Section)
- Capitalizing all the section to Section in all documents;
- Data-Centric storage is the term used in the community. Added reference;
- Changing drawn to figure
- Fact changed to Valid... (Old reference)
- Updating the Data Description Taxonomy Diagram, showing Annotation;
- Adding the description of relating system organization and type of query used on each of them.
- Fixed word agreement
- Additional description about the relationships.


git-svn-id: svn://localhost/net

Taxonomy is defined as the practice and science of classification and uses
the taxonomic units of ``taxa'', the plural form of ``taxon'' \cite{taxonomy}.
Similarly, they are represented by a hierarchical diagram that relates the
main taxonomy to the different taxa and their own related ones, if any.

@@

As delineated in section \ref{sec:sn-data-purpose}, the purpose of the
collected data from sensor devices determines how the collected data is
retrieved and used in the network. In such a fashion, this taxonomy relates to
how the collected data is used, and should be the primary question answered 
during the analysis of a sensor network. The taxa related to this
taxonomy, depicted in Figure \ref{fig:taxonomy-data-purpose}, is described as
follows:

@@

The location where the collected data is stored plays an important role
in the different aspects of the access to the collected data from sensor
devices, as shown in section \ref{sec:sn-storage-locations}. Figure 
\ref{fig:taxonomy-data-location} depicts this taxonomy, which can be
described as follows:

@@

\begin{itemize}
  \item \textbf{Local Storage}: characterizes the place the collected data is
  temporarily located in-memory, or in a secondary storage device on a node
  located in-network;
  \item \textbf{External Storage}: The collected data is usually located in an
  external storage device with a data management system such as a relational
  database. When the collected data occupies different external storage devices, the
  strategy is called \textbf{Data-Centric Storage}, being organized by
  categories based on the collected data keys and values. In other words, what
  characterizes this taxon is the presence of a data partitioning strategy to
  store the collected data.
\end{itemize}

@@

One of the fundamental challenges in the development of a persistence storage
in sensor networks is related to the data model chosen to represent the
collected data. This taxonomy relates to the data model used to design the 
collected data, as drawn in Figure \ref{fig:taxonomy-data-model}, and is 
described as follows.

@@

Particularly related to how the collected data is described, different projects
provide data persistence describing the collected data by using its natural
properties, or associating textual descriptions or keywords for the data. The
way to describe or label data can follow different guidelines such as data provenance
and annotations, discussed in section \ref{sec:sn-data-description}. The description of the
properties should take into account the different values of the nature of the
data and any other description needed. Therefore, the taxa that composes the
Data Description taxonomy, shown in Figure
\ref{fig:taxonomy-data-description}, are described as follows:

@@

\begin{itemize}
  \item \textbf{Provenance-Aware}: when the collected data is categorized based
  on the nature of the data, \textbf{What: Data Identity} is the data type
  that uniquely identifies the collected data from sensors. Similarly, the
  \textbf{When: Time Dimension}, is related to instances of time relative to the
  collected data: \textbf{fact} and \textbf{transaction} times. The former is
  used to identify when the data was collected, whereas the latter when the data was
  transmitted to the data sink where the query processing acts. Finally, the
  \textbf{Where: Data Location} identifies the point where the collected data
  was observed by the sensor device. Usually the latitude and longitude as the
  GPS coordinates;
  \item \textbf{Annotations}: the annotation of the collected data, through the
  use of keywords, full-text description, or any other method that helps
  describe the collected data. It is important to note that this type of data
  does not makeup part of the collected data from sensors. Rather, they
  complement the observations of what was perceived by the sensor devices.
\end{itemize}

@@

The query processing mechanism in sensor networks is directly related to how
and where the collected data is used and stored, as shown in section
\ref{sec:query-process}. The following taxa are related to the
different types of query processing mechanisms and summarized in Figure
\ref{fig:taxonomy-query-mechanism}:

@@

When it comes to the data volume produced during a time frame, different ranges
are reported from different types of sensor networks. Depending on the types
of data produced such as numerical or images, they can be in the orders of
Megabytes (1MB = 1000KB = ~10\superscript{3} bytes), Gigabytes (1GB =
1000\superscript{3} = 10\superscript{9} bytes) and even Terabytes (1TB =
1000\superscript{4} = 10\superscript{12}) to future Petabytes (1PB =
1000\superscript{5} = 10\superscript{15} bytes) as shown in section
\ref{sec:data-load}. As a comparison of what is processed, Google reports to
process 20 Petabytes of data per day in their data centers
\cite{map-reduce-load}. In this way, the following taxa are related to the
different sizes of data produced by a sensor network, summarized in Figure
\ref{fig:taxonomy-data-volume}:

@@

\begin{itemize}
  \item \textbf{Centralized System}: a database system that runs in one single
  centralized host, the focus being the managing of the collected data
  \cite{sn-intro01};
  \item \textbf{Distributed System}: a database system that can be composed of
  more than one node. Usually, the use of the pattern of Master-Slave
  mechanism or data ``Replication" is used to alleviate ``hot spots'' during
  concurrent processes of data collection from the sensor devices and data
  access from the final users. Unique mechanisms can be used to disaggregate 
  the produced data from a concentrated one: the use of database partitioning, 
  as related to the data-centric storage, where each partition is responsible 
  to handle data produced by a given type of sensor, for instance.
\end{itemize}

@@

The taxonomy related to the data model can be linked to any taxon from the
taxon described from the system organization, to data description and query
processing and location of the collected data. In the former, the data of a
relational model is stored using SQL messages.

Similarly, the taxonomy related to the volume of data produced by a sensor
network usually determines the system organization used to handle and process
the collected data. When the volume is small, the use a single server is
required, usually using an external storage device, which relates to the
taxonomy for storage system. However, when the volume is medium to large, the
use of distributed systems is required to cope with high volume of data
processing. Techniques described for data replication is used to decrease the
load in hot spots of the architecture. On the other hand, approaches to
separate the data being written and read can be used such as Load Balancing or
Database Partition, and the use of a Data-Centric storage approach is used.

>>>>

The root of word ``Taxonomy'' is the Greek taxis, meaning 'order' or
'arrangement', and, nomos in Latin, meaning 'law' or 'science'). In
short, taxonomy is defined as the practice and science of classification and
uses the taxonomic units of ``taxa'', the plural form of ``taxon''
\cite{taxonomy}. Similarly, hierarchical diagram is used to relate each of the
taxa to their main taxon.

@@

As delineated in Section \ref{sec:sn-data-purpose}, the purpose of the
collected data from sensor devices determines how the collected data is
retrieved and used in the network. In such a fashion, this taxonomy relates to
how the collected data is used, and should be the primary question answered 
during the analysis of a sensor network. The taxa related to this
taxonomy, depicted in Figure \ref{fig:taxonomy-data-purpose}, is described as
follows:

@@

The location where the collected data is stored plays an important role
in the different aspects of the access to the collected data from sensor
devices, as shown in Section \ref{sec:sn-storage-locations}. Figure 
\ref{fig:taxonomy-data-location} depicts this taxonomy, which can be
described as follows:

@@

\begin{itemize}
  \item \textbf{Local Storage}: characterizes the place the collected data is
  temporarily located in-memory, or in a secondary storage device on a node
  located in-network;
  \item \textbf{External Storage}: The collected data is usually located in an
  external storage device with a data management system such as a relational
  database. When the collected data occupies different external storage devices, the
  strategy is called \textbf{Data-Centric Storage}
  \cite{sn-data-centric-storage}, being organized by categories based on the
  collected data keys and values. In other words, what characterizes this
  taxon is the presence of a data partitioning strategy to store the collected
  data.
\end{itemize}

@@

One of the fundamental challenges in the development of a persistence storage
in sensor networks is related to the data model chosen to represent the
collected data. This taxonomy relates to the data model used to design the 
collected data, as shown in Figure \ref{fig:taxonomy-data-model}, and is 
described as follows.

@@

Particularly related to how the collected data is described, different projects
provide data persistence describing the collected data by using its natural
properties, or associating textual descriptions or keywords for the data. The
way to describe or label data can follow different guidelines such as data provenance
and annotations, discussed in Section \ref{sec:sn-data-description}. The description of the
properties should take into account the different values of the nature of the
data and any other description needed. Therefore, the taxa that compose the
Data Description taxonomy, shown in Figure
\ref{fig:taxonomy-data-description}, are described as follows:

@@

\begin{itemize}
  \item \textbf{Provenance-Aware}: when the collected data is categorized based
  on the nature of the data, \textbf{What: Data Identity} is the data type
  that uniquely identifies the collected data from sensors. Similarly, the
  \textbf{When: Time Dimension}, is related to instances of time relative to the
  collected data: \textbf{valid} and \textbf{transaction} times. The former is
  used to identify when the data was collected, whereas the latter when the data was
  transmitted to the data sink where the query processing acts. Finally, the
  \textbf{Where: Data Location} identifies the point where the collected data
  was observed by the sensor device. Usually the latitude and longitude as the
  GPS coordinates;
  \item \textbf{Annotations}: the annotation of the collected data, through the
  use of keywords, full-text description, or any other method that helps
  describe the collected data. It is important to note that this type of data
  does not makeup part of the collected data from sensors. Rather, they
  complement the observations of what was perceived by the sensor devices.
\end{itemize}

@@

The query processing mechanism in sensor networks is directly related to how
and where the collected data is used and stored, as shown in Section
\ref{sec:query-process}. The following taxa are related to the
different types of query processing mechanisms and summarized in Figure
\ref{fig:taxonomy-query-mechanism}:

@@

When it comes to the data volume produced during a time frame, different ranges
are reported from different types of sensor networks. Depending on the types
of data produced such as numerical or images, they can be in the orders of
Megabytes (1MB = 1000KB = ~10\superscript{3} bytes), Gigabytes (1GB =
1000\superscript{3} = 10\superscript{9} bytes) and even Terabytes (1TB =
1000\superscript{4} = 10\superscript{12}) to future Petabytes (1PB =
1000\superscript{5} = 10\superscript{15} bytes) as shown in Section
\ref{sec:data-load}. As a comparison of what is processed, Google reports to
process 20 Petabytes of data per day in their data centers
\cite{map-reduce-load}. In this way, the following taxa are related to the
different sizes of data produced by a sensor network, summarized in Figure
\ref{fig:taxonomy-data-volume}:

@@

\begin{itemize}
  \item \textbf{Centralized System}: a database system that runs in one single
  centralized host, the focus being the managing of the collected data
  \cite{sn-intro01}. When the system is organized in a centralized system,
  usually a centralized query processing is used to process any request to
  operations on the database management system;
  \item \textbf{Distributed System}: a database system that can be composed of
  more than one node. Usually, the use of the pattern of Master-Slave
  mechanism or data replication \cite{sn-data-center-replication-load-balance} 
  is used to alleviate ``hot spots'' during concurrent processes of data
  collection from the sensor devices and data access from the final users.
  Unique mechanisms can be used to disaggregate the produced data from a
  concentrated one: the use of database partitioning, as related to the
  data-centric storage, where each partition is responsible to handle data
 produced by a given type of sensor, for instance. Finally, it is the type of
 system used to support in-network query processing.
\end{itemize}

@@

The taxonomy related to the data model can be linked to any taxon from the 
system organization because it is used in both centralized and distributed
systems, also associated with the location of the data and, consequently, the
query processing method used. Furthermore, the data description taxon can be
used by any of the data models.

The query processing is directly related to the location of the collected data
and with the system organization taxa. While centralized systems use centralized
query processing methods, in-network query processing is when the collected
data is located in different external storage devices organized in a
data-centric way, usually supported by a distributed system.

Similarly, the taxonomy related to the volume of data produced by a sensor
network usually determines the system organization used to handle and process
the collected data. When the volume is small, the use of a single server is
required, usually using an external storage device, which relates to the
taxonomy for storage systems. However, when the volume is medium to large, the
use of distributed systems is required to cope with high volume of data
processing. Techniques described for data replication is used to decrease the
load in hot spots of the architecture. On the other hand, approaches to
separate the data being written and read can be used such as Load Balancing or
Database Partition, and the use of a Data-Centric storage approach is used.
